Lock-free Skiplist-based Priority Queues \\ Student Paper for Seminar in Algorithms 2013W \\ Technical University of Vienna

### 2014-02-20 paper/paper.tex
Further work on concepts and definitions

>>>>

An area in memory accessed frequently by a large number of processes is said to be \emph{contended}.
Contention is an limiting factor regarding scalability: concurrent reads and writes to the same 
location must be serialized by the cache coherence protocol, and only a single concurrent \ac{CAS} % TODO: Short section on primitives.
can succeed while all others must retry. \emph{Disjoint-access parallelism} is the concept of
spreading such accesses in order to reduce contention as much as possible.

% Priority queues.

Priority queues are an abstract data structure allowing insertion an items with a given priority
and removal of the highest-priority item in logarithmic time. Search trees and Heaps (which are 
flattened representations complete trees such that each node's value is at least as large as
those of both children) are concrete data structures which are usually used to implement sequential
priority queues. However, both require fairly elaborate reorganization after \lstinline|DeleteMin|
and/or \lstinline|Insert| operations, which are especially challenging to achieve in a concurrent
environment.

% SkipLists.

SkipLists \cite{pugh1990skip} have become increasingly popular in concurrent data structures because 
they are both simple and exhibit excellent disjoint-access parallelism. In contrast to
both search trees and Heaps, reorganization is not necessary since SkipLists rely on
randomization for an expected $O(\log n)$ complexity of \lstinline|Insert|, \lstinline|Search|
and \lstinline|Delete| operations.
A SkipList may be visualized as a set of linked lists with corresponding levels. The linked
list at level 0 is the sorted sequence of all objects in the SkipList,
and higher levels provide ``shortcuts''
into the SkipList such that a list of level $i + 1$ contains a subset of the objects in level
$i$. A SkipList node $n$ is said to be of level $i$ if it is in all lists of levels $[0, i]$
and in none of levels $[i + 1, \infty]$. Upon insertion, the new node's level is assigned at random
according to a geometric distribution; deletion simply removes the node.
